---
title: "Using dplyr to fix mistakes"
author: "Marliese"
date: "10/18/2022"
output: html_document
---

Hello! In this homework I will be looking to use dplyr to fix mistakes. This should be very helpful to me based on the last homework lol. I hope it works, but even if it doesn't it will be a learning experience! Okay, lets get into it.

1. First I'm going to set things up.

```{r}
rm(list=ls())
library(ggplot2)
library(dplyr)
library(readr)
library(here)
```

2. Okay now that that is taken care of, lets import the data and run a summary. Dont use stringsAsFactors=TRUE

```{r}
library(readr)
DF <- read.csv(here("Data","drug_resistance.csv"))
View(DF)
```

```{r}
summary(DF)
```

Okay this time use stringsAsFactors=TRUE

```{r}
library(readr)
DF <- read.csv(here("Data","drug_resistance.csv"), stringsAsFactors = TRUE)
View(DF)
```

```{r}
summary(DF)
```

These are different outputs!
W/out stringsAsFactors=TRUE: defines the type of data (character, numerial, etc.) for treatment and tech.name
W/ stringsAsFactors=TRUE: doesn't define the types of data


3. Now use either the function unique() or levels() to examine each of the text-based columns of data. In which variables do you see errors?

```{r}
unique(DF$Treatment)
unique(DF$tech.name)
```
yessss you are an academic weapon! I am so glad that worked!


Both variables have errors. They spelled and capitalized the inputs differently, causing R to think they are different. The Treatments should only have Control, Drug_2, and Drug_3 as options. The tech names should only have Jose, Brian, Mary, and Becca as options.



4. Using the techniques we discussed in class, repair the data set so that there are the appropriate number of choices for each of the character/factor variables.

Okay this is going to be difficult because I wansn't in class, but I will try.


```{r}
no<-which(DF$Treatment=="cantrol"|DF$Treatment=="Control")

DF$Treatment[no]<-"control"


wrong<-which(DF$Treatment=="drug x 2"|DF$Treatment=="drug2")

DF$Treatment[wrong]<-"Drugx2"


ew<-which(DF$Treatment=="drug x 3"|DF$Treatment=="drug3")

DF$Treatment[ew]<-"Drug x 3"


thenerve<-which(DF$tech.name=="brain"|DF$tech.name=="brian")

DF$tech.name[thenerve]<-"Brian"


howdareyou<-which(DF$tech.name=="becca")
DF$tech.name[howdareyou]<-"Becca"

```

thank you so much liv literally saving my life here




