---
title: "Baer_HW11_GSWRCh3Ex1"
author: "Marliese"
date: "10/11/2022"
output: html_document
---

Hello! Welcome to my markdown, in which I will be working through Chapter 3 (not including appendix 3A or 3B) of Getting Started with R.

First I will "set things up"
```{r}
rm(list=ls())
library(dplyr)
library(ggplot2)
library(here)
```



```{r}
DF <- read.csv(here("Data", "compensation (1).csv"), stringsAsFactors=TRUE)
str(DF) #structure
```

See what overall the data looks like.
```{r}
glimpse(DF)
```

Summary Stats!
```{r}
#get summary stats for DF
summary(DF)
```

Now we move on to using dplyr! Lets start w subsetting.

Select a column using select()
```{r}
names(DF) #shows you the name of the columns
select(DF, Fruit) #use the fruit column
```

To select all columns except one:
```{r}
select(DF,-Root) #minus sign followed by a column header will exclude that column from the output
```


Okay, well done! Youre killing it! On to slice(). This function gives rows of data.
```{r}
slice(DF,2) #2 indicates it will give the second row
```


If we want the 2nd row-10th row
```{r}
slice(DF,2:10)
```

for discontinuous sets
```{r}
slice(DF,c(2,3,10))
```
note: it creates new row numbers and doesn't keep the rows they were originally in (in DF)


Nicely done! Moving on to filter(). Before we actually use filter, lets see how it works. How does R interpret ">"
```{r}
with(DF,Fruit>80)
```

filter() uses this! It will sort the data into the categories of true or false and then show you what you want to see.

```{r}
#find rows where it is true that Fruit is >80 return
#them as a data frame
filter(DF,Fruit>80)
```


Select rows w multiple conditions
ie.keep only rows w fruit>80 or less than 20
```{r}
filter(DF,Fruit>80|Fruit<20)
```

Other logical and boolean operations you can use w filter()

```{r}
#to find which rows it is true the conditions are met 
filter(DF,Fruit==80)
#to find which rows it is true the conditions are not met
filter(DF,Fruit!=80)
#to find less than or equal to
filter(DF,Fruit<=80)
#OR: in which rows is it true that fruit is >60 or fruit is <20? it will return all those rows. | = or
filter(DF,Fruit>80|Fruit<20)
#AND: both conditions must be true. this will return rows where it is true that both conditions (from 2 variables) are met. & = and
filter(DF,Fruit>80&Root<2.3)
```


Okay! Now lets make sure we can actually use the subset!

```{r}
lo_hi_fruit<-filter(DF,Fruit>80|Fruit<20)
#now look at it
lo_hi_fruit
```

REMEMBER: to use a subset using filter(), you have to assign the values to a object aka word (as shown above).

Okay, lets move on to transforming.

mutate() is a dplyr function and designates a new column name and the transformation
```{r}
DF<-mutate(DF,logFruit=log(Fruit))
head(DF)
```


Sorting (in this case in order of increasing Fruit production)
```{r}
arrange(DF, Fruit)
```

Using dplyr functions together

```{r}
select(filter(DF,Fruit>80),Root) #filter the data first, take the data frame from filter, and use select to get root column only
```


Oh boy, okay lets get into a concept that is a bit more complicated for me: piping commands!

1. always start with the data frame
```{r}
DF %>% #work w all the DF data
  filter(Fruit>80)%>% #filter it based on the fruit column, getting all rows where it is true that Fruit>80 and then pass this dataframe on to...
  select(Root) #select() and return only the root column as the final data frame!
```


Group summary stats - summarization

method 1: nested, no pipe
```{r}
mean.fruit<-summarise(
  group_by(DF,Grazing),
  meanFruit = mean(Fruit))
```

NOTE: there is no diff between the code directly above and directly below. it is just a formatting preference.

```{r}
mean.fruit<-summarise(group_by(DF,Grazing),meanFruit = mean(Fruit))
```


method 2: pipe, no nesting
```{r}
DF%>%
  group_by(Grazing)%>%
  summarise(meanFruit=mean(Fruit))
```

extending the summarization
```{r}
DF%>%
  group_by(Grazing)%>%
  summarise(
    meanFruit=mean(Fruit),
    sdFruit=sd(Fruit),
    medFruit=median(Fruit))
```

YOU did it!! Yay, well done! Academic weapon!




