---
title: "Baer_Homwork12.Rmd"
author: "Marliese"
date: "10/18/2022"
output: html_document
---

set it up!

```{r}
rm(list=ls())
library(ggplot2)
library(dplyr)
library(here)
```

import dataset
```{r}
library(readr)
Cover <- read.csv(here("Data","donnerville_percent_cover.csv"), stringsAsFactors=TRUE)
View(Cover)
```
I don't know why this is giving me an error in the code but showing up in my environment with no problem. I am just going to keep going because it is in my environment and I can use it so hopefully this won't creat problems later.

UPDATE: I fixed the problem there yay!

GREAT - I DIDN'T GET THE ERROR WHEN I RAN THE CODE!


Examine the dataset!

using str() and glimpse(): these give similar outputs

```{r}
str(Cover)
glimpse (Cover)
```

look at the summary stats

```{r}
summary(Cover)
```

Determine the average percent cover by cover type (bare, bryophytes, vascular plants) for Donnerville 1.

Okay lets try this!

```{r}
mean.percent<-summarise(
  group_by(Cover,Type),
  meanpercent = mean(Percent))
```

Okay, I am having some difficulty. This code almost did what I wanted it to, but I don't think it is only giving me output for site Donerville 1. 

Let me try this.


```{r}
mean.percent<-summarise(
  (filter(Cover,Site=="Donnerville 1"))
  group_by(Cover,Type),
  meanpercent = mean(Percent))
```

Okay, so this isn't really working either. And now my text is in blue? weird.

Now I will try to determine the average percent vascular plant cover across all sites. I am going to try it with piping.

```{r}
Cover %>%
  filter(Type=="Vascular")%>%
  mean(Percent)
```



Okay, so this is not going wicked smoothly. I have been working on this for too long and I am still having issues. I will ask Liv for help and then if we can't figure it out, I will just leave it at this because I don't know why it isn't working.

OKAY - SO IT ISN'T WORKING. I'M GIVING YOU THE POINTS FOR ATTEMPTING IT, BUT WANT TO MAKE SURE YOU CAN DO ALL OF THIS. TAKE A LOOK AT THE KEY.

